---
title: "Sampling Station Exploration"
output: html_document
date: "latest update 2023-05-31"
---

I'm breaking down counts by region to ensure we don't inadvertently leave out a certain area that might be "younger" with whatever time period we determine is enough for long-term analyses.  

Per Melissa, the regions in this spreadsheet are primarily for internal use. And I'm only using them illustratively here, not analytically.  

Also per Melissa, the regions are:  

No. | Region | Reserves |  
:--: | :---- | :---- | 
0 | Northeast | GRB, HUD, NAR, WQB, Wells | 
1 | Mid Atlantic | CBM, CBV, DEL, JAC | 
2 | Southeast | ACE, GTM, NC, NIW, SAP | 
3 | Caribbean | JOB | 
4 | GOM | APA, GND, MAR, RKB, WKB | 
5 | Great Lakes | OWC, LKS | 
6 | West Coast | ELK, HEE, KAC, PDB, SFB, SOS, TJR | 

***  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
mdat <- readr::read_csv(here::here("metadata", "sampling_stations.csv")) %>%
    janitor::clean_names()
```

```{r}
mdat <- mdat %>% 
    select(-row, -lat_long) %>% 
    separate(active_dates, into = c("start_date", "end_date"),
             sep = "-") %>% 
    mutate(station_type = as.character(station_type),
           station_type = case_match(station_type,
                                     "0" ~ "met",
                                     "1" ~ "wq",
                                     "2" ~ "nut"),
           is_swmp = case_match(is_swmp,
                                "P" ~ "Primary",
                                "S" ~ "Secondary"),
           region = as.character(region),
           region = case_when(region == "7" ~ "6",  # 7 is legacy coding and should be 6 
                              .default = region),
           region2 = case_match(region,
                               "0" ~ "0: Northeast",
                               "1" ~ "1: Mid-Atlantic",
                               "2" ~ "2: Southeast",
                               "3" ~ "3: Caribbean",
                               "4" ~ "4: Gulf of Mexico",
                               "5" ~ "5: Great Lakes",
                               "6" ~ "6: West Coast"),   
           state = toupper(state))

active <- mdat %>% 
    filter(status == "Active") %>% 
    mutate(starts = lubridate::decimal_date(lubridate::my(start_date)),
           years_monitored = round(2023.0 - starts, 2))
```

```{r}
ggplot(filter(active, station_type != "nut")) +
    geom_histogram(aes(x = years_monitored, 
                       fill = is_swmp),
                   col = "gray40") +
    geom_vline(xintercept = c(10, 15),
               col = "blue",
               linetype = "dashed") +
    facet_grid(region ~ station_type) +
    labs(title = "WQ and MET stations, by region",
         subtitle = "blue dashed lines at 10y and 15y",
         x = "Years monitored",
         y = "# of stations",
         fill = "") +
    theme_bw()
```

```{r}
ggplot(filter(active, station_type == "nut")) +
    geom_histogram(aes(x = years_monitored, fill = is_swmp),
                   col = "gray40") +
    geom_vline(xintercept = c(10, 15),
               col = "blue",
               linetype = "dashed") +
    facet_wrap(~region2) +
    labs(title = "NUT stations, by region",
         subtitle = "blue dashed lines at 10y and 15y",
         x = "Years monitored",
         y = "# of stations",
         fill = "") +
    theme_bw()
```

```{r}
summ <- active %>% 
    select(-region) %>% 
    rename(region = region2) %>% 
    mutate(yrs = case_when(years_monitored  < 10 ~ "<10",
                           years_monitored >=15 ~ ">15",
                            .default = "10-15")) %>% 
    group_by(station_type, yrs, region) %>% 
    tally()

summ2 <- summ %>% 
    pivot_wider(names_from = yrs,
                values_from = n,
                values_fill = 0) %>% 
    arrange(station_type, region) %>% 
    relocate(`10-15`, .before = `>15`)
```

```{r}
library(kableExtra)

summ2 %>% 
    ungroup() %>% 
    select(-station_type) %>% 
    kbl(align = "lccc",
        caption ="Length of SWMP monitoring by station type and region") %>%
    kable_paper(full_width = F, position = "left") %>%
    add_header_above(c(" " = 1, "Years Monitored" = 3)) %>% 
    pack_rows("MET",
               start_row = 1,
               end_row = 7) %>% 
    pack_rows("NUT",
               start_row = 8,
               end_row = 14) %>% 
    pack_rows("WQ",
               start_row = 15,
               end_row = 21) %>% 
    footnote(general = "Years Monitored calculated by subtracting the starting month/year for each station from the end of 2022.")
```


